---
title: "Mini project"
author: "Group 01: Aditya Samaroo & Yanqing Liu"
date: "01/03/2019"
output:
  html_document:
    highlight: haddock
    number_sections: no
    self_contained: yes
    theme: cerulean
    toc: yes
    toc_depth: 5
    toc_float: yes
  word_document:
    toc: yes
    toc_depth: '5'
  pdf_document:
    toc: yes
    toc_depth: '5'
---
```{r, echo=TRUE, message=FALSE, results='hide', eval=TRUE, warning=F}
library("data.table")
library("dplyr")
library("lubridate")
library("ggplot2")
library("gridExtra")
library("kableExtra")
library("pander")
library("stringr")
library("tm")
library("SnowballC")
library("wordcloud")
library("RColorBrewer")
```

```{r, echo=TRUE, message=FALSE, results='hide', eval=TRUE, warning=F}
accidents <- read.csv("NYPD_Motor_Vehicle_Collisions.csv")
accidents = data.table(accidents) ## 1441945 lines
accidents = na.omit(accidents) ## 986984 lines
```

# World cloud for the first contribution factor of vehicle
```{r, echo=TRUE, message=FALSE, eval=TRUE, warning=F}
data1 = accidents %>% filter(!CONTRIBUTING.FACTOR.VEHICLE.1 %in% 
                               c("Unspecified", "1", "80", "", " "))
reason1 = data.frame(table(data1$CONTRIBUTING.FACTOR.VEHICLE.1))

wordcloud(words = reason1$Var1, freq = reason1$Freq#, type ="file"
          , scale = c(2,0.8), min.freq = 1
          , max.words = Inf, random.order = F, rot.per = 0.1
          , colors = brewer.pal(8, "Dark2"))
```
<a href="#header">back to top</a>

> Please note: Driver inattention or distraction is the most important factor to vehicle collisions.
The second important factor is the Failure to Yield Right-of-way.
Here, the size of word shows the importance.

</br></br></br></br>

# Trend of accidents by date
```{r, echo=TRUE, message=FALSE, eval=TRUE, warning=F}
counted <- accidents %>% group_by(DATE) %>% count(TIME)
counted <- counted %>% mutate(dates = as.Date(DATE, format = "%m/%d/%Y"))
counted_V2 <- counted %>% group_by(dates) %>% summarise(incidents = sum(n))
ggplot(counted_V2, aes(dates, incidents)) + 
  geom_line() + 
  theme_minimal() + 
  xlab("Dates") + ylab("Incidents") + 
  ggtitle("Number of Accidents from 2012 to 2018")
```
<a href="#header">back to top</a>

> Please note: Less accidents recent years than before, normally more accidents during the summer

</br></br>

# Relationship between type of vehicle and accident

## Using first vehicle code type and contribution factor 1
```{r echo=TRUE, message=FALSE, eval=TRUE, warning=F}
type1 <- accidents %>% group_by(VEHICLE.TYPE.CODE.1) %>% 
  count(CONTRIBUTING.FACTOR.VEHICLE.1) %>%
  arrange(desc(n))
type1[type1 == ""] <- NA
type1 <- na.omit(type1)
type1 <- data.frame(type1)
temp1 <- filter(type1, CONTRIBUTING.FACTOR.VEHICLE.1 != "Unspecified")
temp11 <- filter(type1, 
                 CONTRIBUTING.FACTOR.VEHICLE.1 != "Driver Inattention/Distraction" &
                   CONTRIBUTING.FACTOR.VEHICLE.1 != "Unspecified" & n > 50)

p1 <- ggplot(temp11, aes(x = reorder(CONTRIBUTING.FACTOR.VEHICLE.1, n), y = n, 
                         fill = VEHICLE.TYPE.CODE.1)) + 
  geom_bar(stat = "identity") + theme_minimal() +
  theme(axis.text.y = element_text(size = 6), 
        legend.key.width = unit(0.2,"cm"), legend.key.height=unit(0.2,"cm"), 
        legend.position = c(.55, .03), legend.justification = c(0, 0),
        legend.background = element_rect(fill=alpha('white', 0.4)),
        legend.text = element_text(size = 5.5),
        legend.title = element_text(size = 6)) + 
  xlab("Factor") + ylab("Incidents") + 
  ggtitle("Incidents by Factor for Vehicle 1") +
  coord_flip() + 
  guides(fill = guide_legend(ncol = 1, title = "Vehicle Type")) 
p1

temp111 <- filter(type1, 
                  CONTRIBUTING.FACTOR.VEHICLE.1 == "Driver Inattention/Distraction",
                  n > 500)
p11 <- ggplot(temp111, aes(x = reorder(VEHICLE.TYPE.CODE.1, n), y = n)) + 
  geom_bar(stat = "identity") + theme_minimal() + 
  theme(axis.text.y = element_text(size = 8)) + 
  xlab("Vehicle Type") + ylab("Incidents") + 
  ggtitle("Driver Inattention by Vehicle Type") + coord_flip()
p11
```
<a href="#header">back to top</a>

> Passenger vehicle is the type of car who contribute the biggest volume in accidents, especially for driver inattention or distraction.

</br></br>

## Using second vehicle code type and contribution factor 2

```{r echo=TRUE, message=FALSE, eval=TRUE, warning=F}
type2 <- accidents %>% group_by(VEHICLE.TYPE.CODE.2) %>% 
  count(CONTRIBUTING.FACTOR.VEHICLE.2) %>% arrange(desc(n))
type2[type2 == ""] <- NA
type2 <- na.omit(type2)
type2 <- data.frame(type2)
temp2 <- filter(type2, CONTRIBUTING.FACTOR.VEHICLE.2 != "Unspecified" & n > 150)

p2 <- ggplot(temp2, aes(x = reorder(CONTRIBUTING.FACTOR.VEHICLE.2,n), y = n, 
                        fill = VEHICLE.TYPE.CODE.2)) + 
  geom_bar(stat = "identity") + theme_minimal() +
  theme(axis.text.y = element_text(size = 8), 
        legend.key.width = unit(0.2,"cm"), legend.key.height=unit(0.2,"cm"), 
        legend.position = c(.5, .05), legend.justification = c(0, 0),
        legend.background = element_rect(fill=alpha('white', 0.6)),
        legend.text = element_text(size = 5.5),
        legend.title = element_text(size = 6)) + 
  xlab("Factor") + ylab("Incidents") + 
  ggtitle("Incidents by Factor for Vehicle 2") +
  coord_flip() + 
  guides(fill = guide_legend(ncol = 1, title = "Vehicle Type")) 
p2
```
<a href="#header">back to top</a>

> Similar to the first vehicle type code, driver inattention makes up the majority of accident factors

</br></br>

## Using third vehicle code type and contributing factor 3

```{r echo=TRUE, message=FALSE, eval=TRUE, warning=F}
type3 <- accidents %>% group_by(VEHICLE.TYPE.CODE.3) %>% 
  count(CONTRIBUTING.FACTOR.VEHICLE.3) %>% 
  arrange(desc(n))
type3[type3 == ""] <- NA
type3 <- na.omit(type3)
type3 <- data.frame(type3)
temp3 <- filter(type3, CONTRIBUTING.FACTOR.VEHICLE.3 != "Unspecified")

p3 <- ggplot(temp3, aes(x = reorder(CONTRIBUTING.FACTOR.VEHICLE.3,n), y = n, 
                        fill = VEHICLE.TYPE.CODE.3)) + 
  geom_bar(stat = "identity") + theme_minimal() + 
  theme(axis.text.y = element_text(size = 5.5), 
       legend.key.width = unit(0.2,"cm"), legend.key.height=unit(0.2,"cm") , 
       legend.position = c(0.6,.05), legend.justification = c(0, 0),
       legend.background = element_rect(fill=alpha('white', 0.4)),
       legend.text = element_text(size = 5),
       legend.title = element_text(size = 6)) + 
  xlab("Factor") + ylab("Incidents") + 
  ggtitle("Incidents by Factor for Vehicle 3") + coord_flip() + 
  guides(fill = guide_legend(ncol = 1, title = "Vehicle Type"))  
p3
```
<a href="#header">back to top</a>

> Past two vehicles, the third vehicle involved in the accident is usually caused by the other vehicles already involved

</br></br>

## Using fourth vehicle code type and contributing factor 4

```{r echo=TRUE, message=FALSE, eval=TRUE, warning=F}
type4 <- accidents %>% group_by(VEHICLE.TYPE.CODE.4) %>% 
  count(CONTRIBUTING.FACTOR.VEHICLE.4) %>%  arrange(desc(n))
type4[type4 == ""] <- NA
type4 <- na.omit(type4)
type4 <- data.frame(type4)
temp4 <- filter(type4, CONTRIBUTING.FACTOR.VEHICLE.4 != "Unspecified")

p4 <- ggplot(temp4, aes(x = reorder(CONTRIBUTING.FACTOR.VEHICLE.4, n), y = n, 
                        fill = VEHICLE.TYPE.CODE.4)) + 
  geom_bar(stat = "identity") + theme_minimal() +
  theme(axis.text.y = element_text(size = 6), 
        legend.key.width = unit(0.2,"cm"), 
        legend.key.height = unit(0.2,"cm"), 
        legend.position = c(.55, .05), 
        legend.justification = c(0, 0),
        legend.background = element_rect(fill=alpha('white', 0.6)),
        legend.text = element_text(size = 5.5),
        legend.title = element_text(size = 6)) + 
  xlab("Factor") + ylab("Incidents") + ggtitle("Incidents by Factor for Vehicle 4") +
  coord_flip() + 
  guides(fill = guide_legend(ncol = 1, title = "Vehicle Type")) 
p4
```
<a href="#header">back to top</a>

> Other vehicles are usually the cause for the fourth vehicle getting involved in an accident

</br></br>

## Using the last vehicle code type and contribution factor 5

```{r echo=TRUE, message=FALSE, eval=TRUE, warning=F}
type5 <- accidents %>% group_by(VEHICLE.TYPE.CODE.5) %>% 
  count(CONTRIBUTING.FACTOR.VEHICLE.5) %>% arrange(desc(n))
type5[type5 == ""] <- NA
type5 <- na.omit(type5)
type5 <- data.frame(type5)
temp5 <- filter(type5, CONTRIBUTING.FACTOR.VEHICLE.5 != "Unspecified")

p5 <- ggplot(temp5, aes(x = reorder(CONTRIBUTING.FACTOR.VEHICLE.5,n), y = n, 
                        fill = VEHICLE.TYPE.CODE.5)) + 
  geom_bar(stat = "identity") + theme_minimal() +
  theme(axis.text.y = element_text(size = 8), 
        legend.key.width = unit(0.2,"cm"), 
        legend.key.height = unit(0.2,"cm"),
        legend.background = element_rect(fill=alpha('white', 0.6)),
        legend.position = c(.57, .05), 
        legend.justification = c(0, 0),
        legend.text = element_text(size = 6),
        legend.title = element_text(size = 6)) + 
  xlab("Factor") + ylab("Incidents") + 
  ggtitle("Incidents by Factor for Vehicle 5") +
  coord_flip() + 
  guides(fill = guide_legend(ncol = 1, title = "Vehicle Type")) 
p5
```
<a href="#header">back to top</a>

> Other vehicles are usually the cause for the fourth vehicle getting involved in an accident

</br></br></br></br>

# Relationship between time and borough for number of person injured
```{r, echo=TRUE, message=FALSE, eval=TRUE, warning=F}
accidents$hour = as.POSIXlt(as.POSIXlt(paste(Sys.Date(), accidents$TIME)))$hour
accidents$daytime = ifelse(accidents$hour >= 7 & accidents$hour < 12
                           , "Morning"
                           , ifelse(accidents$hour >= 12 & accidents$hour < 18
                                    , "Afternoon"
                                    , ifelse(accidents$hour >= 18 & accidents$hour < 24
                                             , "Evening"
                                             , "Night")))


accidents_daytime <- accidents %>% group_by(BOROUGH, daytime) %>% summarise(nb = sum(NUMBER.OF.PERSONS.INJURED, na.rm = TRUE))
ggplot(accidents_daytime, aes(x = daytime, y = BOROUGH, fill = nb)) + geom_tile() + theme_minimal()
```
<a href="#header">back to top</a>

> Brooklyn is the most risky area during the whole day, Staten Island is quite safe during afternoon. Night is always less dangerous than the other time in a day.

