---
title: "Final Project Markdown"
author: "Aditya Samaroo"
date: "3/27/2019"
output: html_document
---

```{r setup, include=FALSE}
library(readxl)
library(readr)
library(ggplot2)
library(dplyr)
library(dbplyr)
library(purrr)
library(xts)
library(ggthemes)
library(maps)
library(mapdata)
library(leaflet)
library(tidyverse)
library(caret)
library(rpart.plot)
library(knitr)
library(kableExtra)
library(randomForest)
library(e1071)
library(klaR)
library(nnet)
library(pROC)
library(neuralnet)
```

```{r echo=TRUE, message=FALSE, warning=FALSE}
wildfires  <- read_csv("us_fires_1.csv", na = "empty")
wildfires2 <- read_csv("us_fires_2.csv", na = "empty")
wildfires3 <- read_csv("us_fires_3.csv", na = "empty")
wildfires4 <- read_csv("us_fires_4.csv", na = "empty")
wildfires5 <- read_csv("us_fires_5.csv", na = "empty")
wildfires6 <- read_csv("us_fires_6.csv", na = "empty")
wildfires7 <- read_csv("us_fires_7.csv", na = "empty")
fires <- as.data.frame(rbind(wildfires, wildfires2, wildfires3, 
               wildfires4, wildfires5, wildfires6, wildfires7))
head(fires)
set.seed(123)
index <- sample(c(TRUE, FALSE), nrow(fires), replace = TRUE, prob = c(0.2, 0.8))
fires <- fires[index, ]
```
# Preliminary Visualizations
## Missing Values

```{r}
NA_amount <- data.frame(name = c(0), n = c(0), percent = (0))
for (i in 1:length(fires)){
  NA_amount[i,1] <- colnames(fires[i])
  NA_amount[i,2] <- sum(is.na(fires[c(i)]))
  NA_amount[i,3] <- (sum(is.na(fires[c(i)]))/nrow(fires))*100
}
NA_amount <- filter(NA_amount, n > 0)
ggplot(NA_amount, aes(x = name, y = percent)) + 
  geom_bar(stat = "identity", fill = "black") +
  labs(x = "Column Name", y = "Percentage", 
       title = "% of Missing Values in the Dataset by Column Name") +
  theme_calc() +
  theme(axis.text.x = element_text(angle = 50, hjust = 1, size = 8))
```

## By Year
```{r}
fires %>% 
    group_by(fire_year) %>%
    summarize(n_fires = n()) %>%
    ggplot(aes(x = fire_year, y = n_fires/1000)) + 
  geom_bar(stat = "identity", fill = "black") +
  geom_smooth(method = 'lm', se = FALSE, 
              linetype = 'dashed', size = 0.4, color = "red") +
  labs(x = "Year", y = "Number of wildfires (thousands)", 
       title = "US Wildfires by Year") + 
  theme_calc()
```

## By Size
```{r}
fires %>% group_by(gr = cut(fire_size, breaks = c(0, 0.5, 1, 
                                                  2, 3, 4,
                                                  5, 6, 7, Inf ),
                            labels = c("0 - 0.5", "0.5 - 1", "1 - 2", 
                                       "2 - 3", "3 - 4", "4 - 5", 
                                       "5 - 6", "6 - 7", "7 +"))) %>%
  summarise(n_fires = n()) %>% 
  ggplot(aes(x = gr, y = n_fires/1000)) +
  geom_bar(stat = "identity", fill = "black") +
  labs(x = "Fire Size (Acres)", y = "Number of Wildfires (thousands)", 
       title = "Wildfire Sizes 1995 to 2015") + 
  geom_text(aes(label = round(n_fires/1000,1)),
            color = "black", vjust = -0.2, size = 3) +
  theme_calc()
```

### Distribution of the fire sizes
```{r}
ggplot(fires, aes(x = "", y = fire_size)) +
  geom_boxplot() +
  theme_classic()
```

```{r echo=TRUE, message=FALSE, warning=FALSE}
ggplot(fires, aes(x = fire_size)) +
  geom_density(fill = "gray", alpha = 0.5) +
  labs(x = "Fire Size (Acres)", y = "Density", 
       title = "Density Plot of Wildfires by Size") +
  theme_classic()
```

```{r}
summary(fires$fire_size)
```

```{r}
outliers <- boxplot(fires$fire_size,plot = FALSE)$out
fire_out <- fires[-which(fires$fire_size %in% outliers),]
ggplot(fire_out, aes(x = "", y = fire_size)) +
  geom_boxplot() +
  theme_classic()
```

```{r}
summary(fire_out$fire_size)
```

```{r echo=TRUE, message=FALSE, warning=FALSE}
ggplot(fire_out, aes(x = fire_size)) +
  geom_density(fill = "black", alpha = 0.3) +
  labs(x = "Fire Size (Acres)", y = "Density", 
       title = "Density Plot of Wildfires by Size without Outliers") +
  theme_classic()
```

## By Size Class
```{r}
fires %>% group_by(fire_size_class) %>% 
  summarise(n_fires = n()) %>%
  ggplot(aes(x = fire_size_class, y = n_fires/1000)) +
  geom_bar(stat = "identity", fill = "black") +
  labs(x = "Fire Size Class", y = "Number of Wildfires (thousands)", 
       title = "Wildfires by Class Size 1995 - 2015") +
  geom_text(aes(label = round(n_fires/1000,1)), 
            color = "black", vjust = -0.2) +
  theme_calc()
```

## By Cause
```{r}
fires %>%
    group_by(stat_cause_descr) %>%
    summarize(n_fires = n()/1000) %>%
    ggplot(aes(x = reorder(stat_cause_descr, n_fires), y = n_fires)) +
    geom_bar(stat = "identity", fill = "black") + 
    coord_flip() + 
    labs(x = "", y = "Number of fires (thousands)", 
         title = "US Wildfires by Cause 1992 to 2015") +
  theme_calc()
```

## Same Graph Rotated
```{r}
fires %>% group_by(stat_cause_descr) %>%
    summarize(n = n()) %>%
    ggplot(aes(x = reorder(stat_cause_descr, -n), y = n/1000)) + 
    geom_bar(stat = "identity", fill = "black") + 
    labs(x = "", y = "# of instances (thousands)", 
         title = "Wildfire Causes") +
  geom_text(aes(label=(round(n/1000,1))),
            vjust = -0.2, color = "black", size = 3) +
  theme_calc() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
```

## By State
```{r}
state.abb    <- append(state.abb, c("DC", "PR"))
state.name   <- append(state.name, c("District of Columbia", "Puerto Rico"))
fires$region <- map_chr(fires$state, function(x) { 
  tolower(state.name[grep(x,state.abb)]) 
  })
state_map <- map_data("state")
fires %>% 
    select(region) %>%
    group_by(region) %>%
    summarize(n = n()) %>%
    right_join(state_map, by = "region") %>%
    ggplot(aes(x = long, y = lat, group = group, fill = n)) + 
    geom_polygon() + 
    geom_path(color = "white") + 
    scale_fill_continuous(low = "#edb500", 
                          high = "#b32424",
                          name = "Number of Fires") + 
    theme_map() + 
    coord_map("albers", lat0 = 30, lat1 = 40) + 
    ggtitle("US Wildfires from 1995 - 2015") + 
    theme(plot.title = element_text(hjust = 0.5))
```

## Top 10 States
```{r echo=TRUE, message=FALSE, warning=FALSE}
fires %>% 
  group_by(state) %>%
  summarise(n_fires = n()) %>%
  arrange(-n_fires) %>%
  top_n(10) %>%
  ggplot(aes(x = reorder(state, -n_fires), y = n_fires/1000)) +
  geom_bar(stat = "identity", fill ="black") +
  geom_text(aes(label=(round(n_fires/1000,1))),
            vjust = -0.2, color = "black", size = 3) +
  labs(x = "State", y = "Number of Wildfires (thousands)", 
       title = "Top 10 States Where Wildfires Occur") +
  theme_calc() 
```

## Focus on Specific State

```{r}
fires %>% filter(state == "CA") %>%
  group_by(stat_cause_descr) %>%
  summarise(n_fires = n()) %>%
  ggplot(aes(x = reorder(stat_cause_descr, -n_fires), y = n_fires/1000)) +
  geom_bar(stat="identity", fill = "black") +
  labs(x = "", y = "Number of Wildfires", 
       title = "Wildfires by Cause in California") + 
  theme_calc() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
```

```{r}
fires %>% filter(state == "GA") %>%
  group_by(stat_cause_descr) %>%
  summarise(n_fires = n()) %>%
  ggplot(aes(x = reorder(stat_cause_descr, -n_fires), y = n_fires/1000)) +
  geom_bar(stat="identity", fill = "black") +
  labs(x = "", y = "Number of Wildfires", 
       title = "Wildfires by Cause in Georgia") + 
  theme_calc() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
```

```{r}
fires %>% filter(state == "TX") %>%
  group_by(stat_cause_descr) %>%
  summarise(n_fires = n()) %>%
  ggplot(aes(x = reorder(stat_cause_descr, -n_fires), 
             y = n_fires/1000)) +
  geom_bar(stat="identity", fill = "black") +
  labs(x = "", y = "Number of Wildfires", 
       title = "Wildfires by Cause in Texas") + 
  theme_calc() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
```

```{r}
fires %>% filter(state == "TX" | state == "GA" | state == "CA") %>%
  group_by(stat_cause_descr,state) %>%
  summarise(n_fires = n()) %>%
  ggplot(aes(x = reorder(stat_cause_descr, -n_fires), 
             y = n_fires/1000, fill = state)) +
  geom_bar(stat="identity", position = "dodge") +
  labs(x = "", y = "Number of Wildfires (thousandsCr)", 
       title = "Wildfires by Cause in the Top 3 Wildfire States") +
  scale_fill_manual("State", values = c("#a80e0e","#d50101","#e09292")) +
  theme_calc() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
```

## By Cause in Each State 
```{r}
plot_feature_state <- function(cause) {
    fires %>% 
        filter(stat_cause_descr == cause) %>%
        select(region) %>%
        group_by(region) %>%
        summarize(n = n()) %>%
        right_join(state_map, by = "region") %>%
        ggplot(aes(x = long, y = lat, group = group, fill = n)) + 
        geom_polygon() + 
        geom_path(color = "white") + 
        scale_fill_continuous(low = "orange", 
                          high = "darkred",
                          name = "Number of fires") + 
        theme_map() + 
        coord_map("albers", lat0 = 30, lat1 = 40) + 
        ggtitle(paste0("US Wildfires Caused by ", cause, "")) + 
        theme(plot.title = element_text(hjust = 0.5))
}
```

```{r}
plot_feature_state(cause = "Arson")
plot_feature_state(cause = "Campfire")
plot_feature_state(cause = "Children")
plot_feature_state(cause = "Debris Burning")
plot_feature_state(cause = "Equipment Use")
plot_feature_state(cause = "Fireworks")
plot_feature_state(cause = "Lightning")
plot_feature_state(cause = "Miscellaneous")
plot_feature_state(cause = "Missing/Undefined")
plot_feature_state(cause = "Powerline")
plot_feature_state(cause = "Smoking")
plot_feature_state(cause = "Structure")
plot_feature_state(cause = "Railroad")
```

# Data Manipulation
## Time Duration
```{r}
fires_times <- fires[c(20:23,26:28)]
fires_times$dt <- format(as.POSIXct(sprintf("%04d",
                                            fires_times$discovery_time),
                                                 format = "%H%M"),"%H:%M")
fires_times$ct <- format(as.POSIXct(sprintf("%04d", fires_times$cont_time),
                                            format = "%H%M"),"%H:%M")
fires_times$ddt <- as.POSIXct(paste(fires_times$discovery_date,
                                    fires_times$dt),
                              format = "%Y-%m-%d %H:%M")
fires_times$cdt <- as.POSIXct(paste(fires_times$cont_date, 
                                              fires_times$ct),
                                        format = "%Y-%m-%d %H:%M")
fires_times$duration <- difftime(fires_times$cdt, 
                                 fires_times$ddt,
                                 units = "hours")
fires_times$duration <- as.numeric(fires_times$duration)
head(fires_times)
```

```{r echo=TRUE, message=FALSE, warning=FALSE}
fires_times %>% 
  group_by(gr = cut(duration, breaks = c(0, 5, 10, 15, 20, 25, 30, Inf), 
                    labels = c("0 - 5", "5 - 10", "10 - 15",
                               "15 - 20", "20 - 25", "25 - 30", "30 +"))) %>%
  summarise(n_duration = n()) %>% na.omit() %>%
  ggplot(aes(x = gr, y = n_duration/1000)) +
  geom_bar(stat = "identity", fill = "black") +
  labs(x = "Duration (hours)", y = "Number of Wildfires (thousands)",
       title = "Number of Wildfires by Duration") +
  theme_calc()
```

## Correlation Matrix 
```{r}
f_corr  <- cbind(fires[c(20,22,24,29,31,32,33)], fires_times[c(12)])
corrmat <- round(cor(f_corr, use = "complete.obs"), 2)
cormat  <- melt(corrmat)
head(cormat)
```

```{r}
ggplot(cormat, aes(x = Var1, y = Var2, fill = value)) + 
  geom_tile() +
  geom_text(data = cormat, 
            aes(y = Var2, label = value), 
            color = "white") +
  scale_fill_continuous(low = "#e09292",
                        high = "#a80e0e",
                        name = "Correlation") +
  labs(x = "", y = "", title = "Correlation Matrix of Select Variables") +
  theme_calc() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.text.y = element_text(angle = -45, vjust = 1))
```
# Classification algorithms
## Simple decision tree using one feature
```{r}
fires$stat_cause_descr <- as.factor(fires$stat_cause_descr)

set.seed(123)
train_index <- sample(c(1:dim(fires)[1]), dim(fires)[1]*0.6)

x_train <- as.data.frame(fires[train_index, features])
colnames(x_train) <- "fire_size"
y_train <- fires$stat_cause_descr[train_index]
x_test <- as.data.frame(fires[-train_index, features])
colnames(x_test) <- "fire_size"
y_test <- fires$stat_cause_descr[-train_index]
```
# Naive Rule
```{r}
preds <- rep('Debris Burning', length(y_test))

test_set_acc <- round(sum(y_test == preds)/length(preds), 4)
print(paste(c("Accuracy:" , test_set_acc)))
```

# Decision Tree
```{r}
# create the training control object.
tr_control <- trainControl(method = 'cv', number = 3)

features <- c('fire_year', 'fire_size', 'discovery_doy', 'latitude', 'longitude')

x_train <- as.data.frame(fires[train_index, features])
colnames(x_train) <- c("fire_year","fire_size","discovery_doy","latitude","longitude")
y_train <- fires$stat_cause_descr[train_index]

x_test <- as.data.frame(fires[-train_index, features])
colnames(x_test) <- c("fire_year","fire_size","discovery_doy","latitude","longitude")
y_test <- fires$stat_cause_descr[-train_index]
```
# Train the decision tree model
```{r}
set.seed(123)
dtree <- train(x = x_train,
               y = y_train,
               method = 'rpart',
               tuneLength = 8,
               trControl = tr_control)
```
# Tree Plot
```{r}
rpart.plot(dtree$finalModel)
```
# make predictions using test set
```{r}
preds <- predict(dtree, newdata = x_test)
```
# calculate accuracy on test set
```{r}
test_set_acc <- sum(y_test == preds)/length(preds)
print(paste(c("Accuracy:" , round(test_set_acc, 4))))
```
# Confusion Matrix
```{r}
confusionMatrix(y_test, preds)$table %>%
    prop.table(margin = 1) %>%
    as.data.frame.matrix() %>%
    rownames_to_column(var = 'actual') %>%
    gather(key = 'prediction', value = 'freq',-actual) %>%
    ggplot(aes(x = actual, y = prediction, fill = freq)) +
    geom_tile() +
    geom_text(aes(label = round(freq, 2)), size = 3, color = 'gray20') + 
    scale_fill_gradient(low = 'yellow', high = 'red', 
                        limits = c(0,1), name = 'Relative Frequency') +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    ggtitle('Confusion Matrix - Decision Tree')
```
# Random Forest
```{r}
set.seed(123)
rfmodel <- train(x = x_train,
                 y = y_train,
                 method = 'rf',
                 tuneLength = 3,
                 ntree = 100,
                 trControl = tr_control)
rfmodel
```
# make predictions using test set
```{r}
preds <- predict(rfmodel, newdata = x_test)
```
# calculate accuracy on test set
```{r}
test_set_acc <- sum(y_test == preds)/length(preds)
print(paste(c("Accuracy:" , round(test_set_acc, 4))))
```
## Confusion Matrix
```{r}
confusionMatrix(y_test, preds)$table %>%
    prop.table(margin = 1) %>%
    as.data.frame.matrix() %>%
    rownames_to_column(var = 'actual') %>%
    gather(key = 'prediction', value = 'freq',-actual) %>%
    ggplot(aes(x = actual, y = prediction, fill = freq)) +
    geom_tile() +
    geom_text(aes(label = round(freq, 2)), size = 3, color = 'gray20') + 
    scale_fill_gradient(low = 'yellow', high = 'red', limits = c(0,1), name = 'Relative Frequency') + 
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    ggtitle('Confusion Matrix - Random Forest (100 Trees)')
```
## Multiclass ROC area under the curve
```{r}
prob <- predict(rfmodel, newdata= x_test, type = "prob")
r <- multiclass.roc(y_test, prob)
r
```
# Boosting
```{r}
fires$burn_time <- fires$cont_date - fires$discovery_date

features <- c('fire_year', 'fire_size', 'discovery_doy', 'latitude', 'longitude','burn_time')

x_train <- as.data.frame(fires[train_index, features])
colnames(x_train) <- c("fire_year","fire_size","discovery_doy","latitude","longitude","burn_time")
y_train <- fires$stat_cause_descr[train_index]

x_test <- as.data.frame(fires[test_index, features])
colnames(x_test) <- c("fire_year","fire_size","discovery_doy","latitude","longitude","burn_time")
y_test <- fires$stat_cause_descr[test_index]
```
# knn
```{r}
set.seed(123)
knn <- train(  x = x_train,
               y = y_train,
               method = 'knn',
               tuneLength = 8,
               trControl = tr_control,
               preProcess = c("center", "scale"))

knn
```
## make predictions using test set
```{r}
preds <- predict(knn, newdata = x_test)
```
## calculate accuracy on test set
```{r}
test_set_acc <- sum(y_test == preds)/length(preds)
print(paste(c("Accuracy:" , round(test_set_acc, 4))))
```
## Confusion Matrix
```{r}
confusionMatrix(y_test, preds)$table %>%
    prop.table(margin = 1) %>%
    as.data.frame.matrix() %>%
    rownames_to_column(var = 'actual') %>%
    gather(key = 'prediction', value = 'freq',-actual) %>%
    ggplot(aes(x = actual, y = prediction, fill = freq)) +
    geom_tile() +
    geom_text(aes(label = round(freq, 2)), size = 3, color = 'gray20') + 
    scale_fill_gradient(low = 'yellow', high = 'red', limits = c(0,1), name = 'Relative Frequency') +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    ggtitle('Confusion Matrix - K Nearest Neighbours')
```
## Multiclass ROC area under the curve
```{r}
prob <- predict(knn, newdata= x_test, type = "prob")
r <- multiclass.roc(y_test, prob)
r
```
# Naive Bayes Classifier
```{r}
set.seed(123)
nb <- train(  x = x_train,
               y = y_train,
               method = 'nb',
               tuneLength = 8,
               trControl = tr_control,
               preProcess = c("center", "scale"))
nb
```
## make predictions using test set
```{r}
preds <- predict(nb, newdata = x_test)
```
## calculate accuracy on test set
```{r}
test_set_acc <- sum(y_test == preds)/length(preds)
print(paste(c("Accuracy:" , round(test_set_acc, 4))))
```
## Confusion Matrix
```{r}
confusionMatrix(y_test, preds)$table %>%
    prop.table(margin = 1) %>%
    as.data.frame.matrix() %>%
    rownames_to_column(var = 'actual') %>%
    gather(key = 'prediction', value = 'freq',-actual) %>%
    ggplot(aes(x = actual, y = prediction, fill = freq)) +
    geom_tile() +
    geom_text(aes(label = round(freq, 2)), size = 3, color = 'gray20') + 
    scale_fill_gradient(low = 'yellow', high = 'red', limits = c(0,1), name = 'Relative Frequency') +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    ggtitle('Confusion Matrix - Naive Bayes Classifier')
```
## Multiclass ROC area under the curve
```{r}
prob <- predict(nb, newdata= x_test, type = "prob")
r <- multiclass.roc(y_test, prob)
r
```
# Logistic Regression 
```{r}
set.seed(123)
log_reg <- train(  x = x_train,
               y = y_train,
               method = 'multinom',
               #tuneLength = 8,
               trControl = tr_control
               #preProcess = c("center", "scale"
               )
log_reg
```
## make predictions using test set
```{r}
preds <- predict(nb, newdata = x_test)
```
## calculate accuracy on test set
```{r}
test_set_acc <- sum(y_test == preds)/length(preds)
print(paste(c("Accuracy:" , round(test_set_acc, 4))))
```
## Confusion Matrix
```{r}
confusionMatrix(y_test, preds)$table %>%
    prop.table(margin = 1) %>%
    as.data.frame.matrix() %>%
    rownames_to_column(var = 'actual') %>%
    gather(key = 'prediction', value = 'freq',-actual) %>%
    ggplot(aes(x = actual, y = prediction, fill = freq)) +
    geom_tile() +
    geom_text(aes(label = round(freq, 2)), size = 3, color = 'gray20') + 
    scale_fill_gradient(low = 'yellow', high = 'red', limits = c(0,1), name = 'Relative Frequency') +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    ggtitle('Confusion Matrix - Multinomial Regression')
```
## Multiclass ROC area under the curve
```{r}
prob <- predict(log_reg, newdata= x_test, type = "prob")
r <- multiclass.roc(y_test, prob)
r
```
# Linear Discriminant Analysis
```{r}
set.seed(123)
lda <- train(  x = x_train,
               y = y_train,
               method = 'lda',
               #tuneLength = 8,
               trControl = tr_control
               #preProcess = c("center", "scale"
               )
lda 
```
## make predictions using test set
```{r}
preds <- predict(lda, newdata = x_test)
```
## calculate accuracy on test set
```{r}
test_set_acc <- sum(y_test == preds)/length(preds)
print(paste(c("Accuracy:" , round(test_set_acc, 4))))
```
## Confusion Matrix
```{r}
confusionMatrix(y_test, preds)$table %>%
    prop.table(margin = 1) %>%
    as.data.frame.matrix() %>%
    rownames_to_column(var = 'actual') %>%
    gather(key = 'prediction', value = 'freq',-actual) %>%
    ggplot(aes(x = actual, y = prediction, fill = freq)) +
    geom_tile() +
    geom_text(aes(label = round(freq, 2)), size = 3, color = 'gray20') + 
    scale_fill_gradient(low = 'yellow', high = 'red', limits = c(0,1), name = 'Relative Frequency') +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    ggtitle('Confusion Matrix - Linear Discriminant Analysis')
```
## Multiclass ROC area under the curve
```{r}
prob <- predict(lda, newdata= x_test, type = "prob")
r <- multiclass.roc(y_test, prob)
r
```
# Neural Network
```{r}
set.seed(123)
nn <- train(x = x_train, 
            y = y_train, 
            method = 'nnet', 
            #linear.output = F,
            trControl = tr_control
            )
```
## make predictions using test set
```{r}
preds <- predict(nn, newdata = x_test)
```
## calculate accuracy on test set
```{r}
test_set_acc <- sum(y_test == preds)/length(preds)
print(paste(c("Accuracy:" , round(test_set_acc, 4))))
```

## Confusion Matrix
```{r}
confusionMatrix(y_test, preds)$table %>%
    prop.table(margin = 1) %>%
    as.data.frame.matrix() %>%
    rownames_to_column(var = 'actual') %>%
    gather(key = 'prediction', value = 'freq',-actual) %>%
    ggplot(aes(x = actual, y = prediction, fill = freq)) +
    geom_tile() +
    geom_text(aes(label = round(freq, 2)), size = 3, color = 'gray20') + 
    scale_fill_gradient(low = 'yellow', high = 'red', limits = c(0,1), name = 'Relative Frequency') +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    ggtitle('Confusion Matrix - Neural Network')
```
## Multiclass ROC area under the curve
```{r}
prob <- predict(nn, newdata= x_test, type = "prob")
r <- multiclass.roc(y_test, prob)
r
```
## Reducing the number of categories in the output variable from 13 to 4
```{r}
y_train <- ifelse(y_train == "Lightning", 0, ifelse(y_train == "Structure", 1, ifelse(y_train == "Fireworks", 2, ifelse(y_train == "Powerline", 3, ifelse(y_train == "Railroad", 4, ifelse(y_train == "Smoking", 5, ifelse(y_train == "Children", 6, ifelse(y_train =="Campfire", 7, ifelse(y_train == "Equipment Use", 8, ifelse(y_train == "Debris Burning", 9, ifelse(y_train == "Arson", 10, ifelse(y_train == "Miscellaneous", 11,12))))))))))))
y_train <- as.factor(ifelse(y_train == 0, "Natural", ifelse(y_train == 1 | y_train == 2 | y_train == 3 | y_train == 4 | y_train == 5 | y_train == 6 | y_train == 7 | y_train == 8 | y_train == 9, "Accidental", ifelse(y_train == 10, "Malicious", "Other"))))

y_test <- ifelse(y_test == "Lightning", 0, ifelse(y_test == "Structure", 1, ifelse(y_test == "Fireworks", 2, ifelse(y_test == "Powerline", 3, ifelse(y_test == "Railroad", 4, ifelse(y_test == "Smoking", 5, ifelse(y_test == "Children", 6, ifelse(y_test =="Campfire", 7, ifelse(y_test == "Equipment Use", 8, ifelse(y_test == "Debris Burning", 9, ifelse(y_test == "Arson", 10, ifelse(y_test == "Miscellaneous", 11,12))))))))))))
y_test <- as.factor(ifelse(y_test == 0, "Natural", ifelse(y_test == 1 | y_test == 2 | y_test == 3 | y_test == 4 | y_test == 5 | y_test == 6 | y_test == 7 | y_test == 8 | y_test == 9, "Accidental", ifelse(y_test == 10, "Malicious", "Other"))))

```
# Random Forest(Reduced Categories)
```{r}
set.seed(123)
rfmodel <- train(x = x_train,
                 y = y_train,
                 method = 'rf',
                 tuneLength = 3,
                 ntree = 100,
                 trControl = tr_control)
```
## Make predictions using test set
```{r}
preds <- predict(rfmodel, newdata = x_test)
```
## calculate accuracy on test set
```{r}
test_set_acc <- sum(y_test == preds)/length(preds)
print(paste(c("Accuracy:" , round(test_set_acc, 4))))
print(rfmodel$resample)
```
## Confusion Matrix
```{r}
confusionMatrix(y_test, preds)$table %>%
    prop.table(margin = 1) %>%
    as.data.frame.matrix() %>%
    rownames_to_column(var = 'actual') %>%
    gather(key = 'prediction', value = 'freq',-actual) %>%
    ggplot(aes(x = actual, y = prediction, fill = freq)) +
    geom_tile() +
    geom_text(aes(label = round(freq, 2)), size = 3, color = 'gray20') + 
    scale_fill_gradient(low = 'yellow', high = 'red', 
                        limits = c(0,1), name = 'Relative Frequency') +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    ggtitle('Confusion Matrix - Random Forest(Only 4 Casses)')
```